Nathan Greenslit Data Analysis Project

### 2023-04-05 code/analysis_code/stats/3_uni_variate.qmd
edits

```{r, include = FALSE}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
```

@@

@@

### RMSE and R2 to test model performance

@@

### AIC to assess Performance

@@

### Plot Residuals

@@

>>>>

```{r, include = FALSE}
library(tidyverse)
library(here)
library(Hmisc) #For lag function
library(tidymodels) #For modeling
library(ggpmisc) #For R2 equation on graph
library(dotwhisker) #Fpr visualization
library(ggpubr) #For plot stacking
```

@@

## Null Model:

#### 5-Fold Cross Validation

```{r}
fold_bo <- vfold_cv(bo2, v = 5, repeats = 5) #Data was too small to stratify (<20)
fold_c2 <- vfold_cv(c22, v = 5, repeats = 5) #Data was too small to stratify (<20)
fold_rd <- vfold_cv(rd2, v = 5, repeats = 5) #Data was too small to stratify (<20)
```

#### Create Null Model

```{r}
#Create Null Recipe for Growth at all three sites
null_bo<- recipe(copies_mL ~ 1, data = bo2) 
null_c2<- recipe(copies_mL ~ 1, data = c22)
null_rd<- recipe(copies_mL ~ 1, data = rd2)

#Workflow that adds recipe to model for three sites
null_wflow_bo<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_bo)

null_wflow_c2<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_c2)

null_wflow_rd<- 
  workflow() %>%
  add_model(lm_mod) %>%
  add_recipe(null_rd)

#Use workflow to fit model to  data set
null_fit_bo<-
  fit_resamples(null_wflow_bo, resamples = fold_bo)

null_fit_c2<-
  fit_resamples(null_wflow_c2, resamples = fold_c2)

null_fit_rd<-
  fit_resamples(null_wflow_rd, resamples = fold_rd)
```

#### Get Metrics for Null

##### RMSE

```{r}
null_met_bo <- 
  collect_metrics(null_fit_bo)

null_met_c2 <- 
  collect_metrics(null_fit_c2)

null_met_rd <- 
  collect_metrics(null_fit_rd)

null_met_bo
null_met_c2
null_met_rd
```

##### AIC

```{r}
bo_null_mod<- 
  lm(copies_mL ~ 1, data = bo2)
AIC(bo_null_mod)

c2_null_mod<- 
  lm(copies_mL ~ 1, data = c22)
AIC(c2_null_mod)

rd_null_mod<- 
  lm(copies_mL ~ 1, data = rd2)
AIC(rd_null_mod)
```

We will come back to these metrics once we run the models on our actual data.

## Actual Data for Model

@@

#### Metrics

##### RMSE and R2

@@

##### AIC

@@

#### Plot Residuals

@@

## Conclusion

| Site | Model Type | RMSE     | AIC      |
|------|------------|----------|----------|
| BO   | Null       | 20472.82 | 275.9331 |
| BO   | Actual     | 14422.91 | 269.8923 |
| C2   | Null       | 8472.898 | 253/1876 |
| C2   | Actual     | 5191.22  | 245.3679 |
| RD   | Null       | 50386.56 | 296.9821 |
| RD   | Actual     | 36132.11 | 291.933  |

AIC and RMSE are lower (better performance) than the null.

<!--# Why are these so high? -->

### 2023-04-05 products/manuscript/Manuscript.qmd
edits

@@

3.  Looking at my statistics/models, is there anything that you think is strikingly missing?

@@

The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986]. Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico (Fig. 1). These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each. Dust aerosols can harbor a wide diversity of bacteria [@kelloggCharacterizationAerosolizedBacteria2004], fungi [@ramirez-camejoFungalDiversitySahara2022], virus-like particles [@griffinAfricanDesertDust], and minerals [@formenti2003]. These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron [@millsIronPhosphorusColimit2004 ; @savoieNonseasaltSulfateNitrate1989 ; @graham1982] . This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups [@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].

@@

Within the *Vibrionaceae* family, there are at least 12 species that are commonly known as human pathogens. Illness can be induced through either a foodborne route via the consumption of raw or uncooked seafood and contaminated drinking water, or through nonfoodborne routes such as wound exposure while swimming. The most common illnesses caused by *Vibrio* include self-limiting diarrhea and cholera (*V. cholerae*), shellfish-induced gastroenteritis (*V. parahaemolyticus, V. vulnificus*), extreme cases of necrotizing fasciitis and septicemia (*V. vulnificus*), and wound and ear infections  (*V. alginolyticus*, *V. vulnificus*) [@baker-austinVibrioSppInfections2018]. Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V. tubiashii, V. coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V. coralliilyticus*) [@richardsMortalitiesEasternPacific2015 ; @rosenbergVibrioShiloiOculina2004; @ben-haimVibrioCoralliilyticusSp2003], and mortality in important aquaculture species such as penaeid shrimp (*V. parahaemolyticus*) and sea breams (*V. harveyi)* [@tranDeterminationInfectiousNature2013 ; @haldarIdentificationVibrioHarveyi2010] among others.

>>>>

        -   Why are my RMSE and AIC so high?

@@

3.  Ultimately, I feel as though I am missing some "oomph"when it comes to my statistical analysis. The linear models seem tadequate, but I know that there are multiple variables at play at these sites (and we have the nutrient data for this). I am open to any suggestions that may diversify and expand my analyses!

@@

The Sahara Desert is a significant source of atmospheric dust, eliciting an estimated one billion metric tons each year [@dalmeidaModelSaharanDust1986\]. Annually, plumes of this dust travel across the Atlantic via easterly trade winds to be deposited in the surface waters of the Atlantic, Caribbean, and Gulf of Mexico (Fig. 1). These dust events are highly episodic, occurring 3- 4 times a year typically in the summer months and lasting 3-5 days each. Dust aerosols can harbor a wide diversity of bacteria \[@kelloggCharacterizationAerosolizedBacteria2004\], fungi \[@ramirez-camejoFungalDiversitySahara2022\], virus-like particles \[@griffinAfricanDesertDust\], and minerals \[@formenti2003\]. These aerosols also serve as a significant source of macro and micronutrients such as phosphate, nitrate, and iron \[@millsIronPhosphorusColimit2004; @savoieNonseasaltSulfateNitrate1989 ; @graham1982\] . This addition of otherwise limiting resources can lead to rapid and potentially harmful blooms of certain microbial groups \[@lenesSaharanDustPhosphatic2008; @westrichSaharanDustNutrients2016a ; @westrichSaharanDustNutrients2016a].

@@

Within the *Vibrionaceae* family, there are at least 12 species that are commonly known as human pathogens. Illness can be induced through either a foodborne route via the consumption of raw or uncooked seafood and contaminated drinking water, or through nonfoodborne routes such as wound exposure while swimming. The most common illnesses caused by *Vibrio* include self-limiting diarrhea and cholera (*V. cholerae*), shellfish-induced gastroenteritis (*V. parahaemolyticus, V. vulnificus*), extreme cases of necrotizing fasciitis and septicemia (*V. vulnificus*), and wound and ear infections  (*V. alginolyticus*, *V. vulnificus*) [@baker-austinVibrioSppInfections2018\]. Some species also pose a threat to marine health, inducing mortality in oyster and clam larvae (*V. tubiashii, V. coralliilyticus*), bacterial bleaching and rapid tissue loss in corals (*V. mediterranei, V. coralliilyticus*) \[@richardsMortalitiesEasternPacific2015; @rosenbergVibrioShiloiOculina2004; @ben-haimVibrioCoralliilyticusSp2003\], and mortality in important aquaculture species such as penaeid shrimp (*V. parahaemolyticus*) and sea breams (*V. harveyi)* \[@tranDeterminationInfectiousNature2013; @haldarIdentificationVibrioHarveyi2010] among others.
